---
title: "p8105_hw2_ll3344"
author: "Lusha Liang"
date: "9/24/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read in the Mr. Trashwheel dataset.

```{r trashwheel}
trashwheel_df = 
  read_xlsx("./data/trash_wheel.xlsx", 
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read in the precipitation data for 2017 and 2018. 

```{r precip}
precip_2018 = 
  read_excel(
    "./data/trash_wheel.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/trash_wheel.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation from 2017 and 2018. 

```{r combine}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

precip_df = 
  left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash (such as sports balls, plastic bottles, and cigarette butts). There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data. 

* The total amount of precipitation in 2018 was `r precip_2018 %>% pull(total) %>% sum()` inches. 
* The median number of sports balls in a dumpster in 2017 was `r trashwheel_df %>% filter(year == "2017") %>% pull(sports_balls) %>% median()`. 

## Problem 2

Read in the NYC transit data. 

```{r}
transit_df =
  read_csv("./data/transit.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>%
  mutate(entry = ifelse(entry == "YES", "true", "false"),
         entry = as.logical(entry))
```

The NYC transit dataset contains information about characteristics of entrances and exits of each subway station in NYC. Thus far we have subsetted a portion of the dataset focusing on subway lines, routes, and information about entrances. The resulting data set has `r nrow(transit_df)` rows and `r ncol(transit_df)`columns. Currently the data is not tidy in that the route variable encompasses both route name and route number. 

Determine the number of stations using the distinct function. 

```{r}
distinct_transit_df =
  transit_df %>% 
  distinct(station_name, .keep_all = TRUE)
```

There are `r nrow(distinct_transit_df)` distinct stations. 

Determine the number of ADA compliant stations using the filter function.

```{r}
ada_station_df =
  distinct_transit_df %>%
  filter(ada == "TRUE")
```

`r nrow(ada_station_df)` stations are ADA compliant. 

Determine what proportion of station entrances/exits without vending allow entrance.

```{r}
without_vending_df = 
  transit_df %>% 
  filter(vending == "NO")

without_vending_entry_df =
  transit_df %>%
  filter(vending == "NO", entry == "TRUE")
```

* The number of station entrances/exits without vending that allow entrance is `r nrow(without_vending_entry_df)`.
* The total number of stations entrances/exits without vending is `r nrow(without_vending_df)`. 
* The proportion of station entrances/exits without vending that allow entry out of all stations without vending is `r nrow(without_vending_entry_df)/nrow(without_vending_df)`. 

Reformat data so they are tidier! More explicitly, reoconfigure the route variable to two distinct variables, one that contains route number and one that contains route name. 

```{r}
tidy_transit_df = 
  distinct_transit_df %>%
  mutate(across(route1:route11, as.character)) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

How many distinct stations serve the A train?

```{r}
a_train_df =
  tidy_transit_df %>%
  filter(route_name == "A")
```

`r nrow(a_train_df)` stations serve the A train. Of the stations that serve the A train, `r a_train_df %>% filter(ada == "TRUE") %>% nrow()` of them are ADA compliant. 

## Problem 3


