---
title: "p8105_hw2_ll3344"
author: "Lusha Liang"
date: "9/24/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read in the Mr. Trashwheel dataset.

```{r trashwheel}
trashwheel_df = 
  read_xlsx("./data/trash_wheel.xlsx", 
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read in the precipitation data for 2017 and 2018. 

```{r precip}
precip_2018 = 
  read_excel(
    "./data/trash_wheel.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/trash_wheel.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation from 2017 and 2018. 

```{r combine}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

precip_df = 
  left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash (such as sports balls, plastic bottles, and cigarette butts). There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data. The total amount of precipitation in 2018 was `r precip_2018 %>% pull(total) %>% sum()` inches. The median number of sports balls in a dumpster in 2017 was `r trashwheel_df %>% filter(year == "2017") %>% pull(sports_balls) %>% median()`. 

