---
title: "p8105_hw2_ll3344"
author: "Lusha Liang"
date: "9/24/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read in the Mr. Trashwheel dataset.

```{r trashwheel}
trashwheel_df = 
  read_xlsx("./data/trash_wheel.xlsx", 
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read in the precipitation data for 2017 and 2018. 

```{r precip}
precip_2018 = 
  read_excel(
    "./data/trash_wheel.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/trash_wheel.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation from 2017 and 2018. 

```{r combine}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

precip_df = 
  left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the harbor, the trashwheel collects that trash, and stores it in a dumpster. The trashwheel_df dataset contains information on year, month, and trash collected, including some specific kinds of trash (such as sports balls, plastic bottles, and cigarette butts). There are a total of `r nrow(trashwheel_df)` rows in our final trashwheel dataset. Additional data sheets include monthly precipitation data. There are a total of `r nrow(precip_df)` observations in the final precipitation dataset which includes precipitation in inches by month for the years 2017 and 2018. 

* The total amount of precipitation in 2018 was `r precip_2018 %>% pull(total) %>% sum()` inches. 
* The median number of sports balls in a dumpster in 2017 was `r trashwheel_df %>% filter(year == "2017") %>% pull(sports_balls) %>% median()`. 

## Problem 2

Read in the NYC transit data. 

```{r}
transit_df =
  read_csv("./data/transit.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>%
  mutate(entry = ifelse(entry == "YES", "true", "false"),
         entry = as.logical(entry))
```

The NYC transit dataset contains information about characteristics of entrances and exits of each subway station in NYC. This includes information on which lines run from each station, whether entrances/exits are ADA compliant, and so forth. Thus far we have subsetted a portion of the dataset focusing on subway lines, routes, and information about entrances. We have also converted entry information from Yes/No format to logical (True/False). The resulting data set has `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. Currently the data is not tidy in that the route variable encompasses both route name and route number. 

Since the data contain multiple rows for each distinct station, we will now determine the number of stations using the distinct function. 

```{r}
distinct_transit_df =
  transit_df %>% 
  distinct(station_name, .keep_all = TRUE)
```

* There are `r nrow(distinct_transit_df)` distinct stations. 
* Of those stations, `r distinct_transit_df %>% filter(ada == "TRUE") %>% nrow()` are ADA compliant.

Determine what proportion of station entrances/exits without vending allow entrance.

```{r}
without_vending_df = 
  transit_df %>% 
  filter(vending == "NO")

without_vending_entry_df =
  without_vending_df %>%
  filter(entry == "TRUE")
```

* The number of station entrances/exits without vending that allow entrance is `r nrow(without_vending_entry_df)`.
* The total number of stations entrances/exits without vending is `r nrow(without_vending_df)`. 
* The proportion of station entrances/exits without vending that allow entry out of all stations without vending is `r nrow(without_vending_entry_df)/nrow(without_vending_df)`. 

Now we will reformat data so they are tidier! More explicitly, we will reoconfigure the route variable to two distinct variables: one that contains route number and one that contains route name. 

```{r}
tidy_transit_df = 
  distinct_transit_df %>%
  mutate(across(route1:route11, as.character)) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

`r tidy_transit_df %>% filter(route_name == "A") %>% nrow()` distinct stations serve the A train. 

```{r}
a_train_df =
  tidy_transit_df %>%
  filter(route_name == "A")
```

Of the stations that serve the A train, `r a_train_df %>% filter(ada == "TRUE") %>% nrow()` are ADA compliant. 

## Problem 3

Read in the FiveThirtyEight data. 

```{r}
pols_df =
  read_csv("./data/fivethirtyeight/pols-month.csv")

unemployment_df =
  read_csv("./data/fivethirtyeight/unemployment.csv")

snp_df =
  read_csv("./data/fivethirtyeight/snp.csv")
```

Clean pols_month data.

```{r}
pols_df = 
  pols_df %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(across(year:day, as.numeric))

pols_df = left_join(pols_df, month_df, by = "month")  

pols_df =
  pols_df %>%
  relocate(year, month_name) %>%
  mutate(president = ifelse(prez_gop == 0, "dem", "gop")) %>%
  arrange(year, month) %>%
  select(-c(month, day, prez_gop, prez_dem)) %>%
  rename(month = month_name)
```
NB: Interestingly, prez_gop has values of 0 for Democrat and 1 for Republican/GOP, but also a few values of 2 between August through December of 1974 which correspond to the transition of the presidency from Richard Nixon to Gerald Ford following Watergate and Nixon's subsequent impeachment. However, since Gerald Ford was a Republican as well, the above code specifying that any value but "0" in the prez_gop column should be coded as "gop" is still correct and produces "gop" in the months between August through December of 1974. 

Clean snp data. 

```{r}
snp_df = 
  snp_df %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(across(month:year, as.numeric))

snp_df = left_join(snp_df, month_df, by = "month")  

snp_df =
  snp_df %>%
  relocate(year, month_name) %>% 
  arrange(year, month) %>%
  select(-c(month, day)) %>%
  rename(month = month_name)
```

Tidy the unemployment data. 

```{r}
unemployment_df = 
  unemployment_df %>%
  rename(year = Year, January = Jan, February = Feb, March = Mar, April = Apr, June = Jun, July = Jul,
         August = Aug, September = Sep, October = Oct, November = Nov, December = Dec) %>%
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "unemployment"
  ) 
```

Join snp with pols. 

```{r}
snp_pols_df = left_join(pols_df, snp_df, by = c("month", "year")) 
```

Join unemployment with the result. 

```{r}
complete_df = left_join(snp_pols_df, unemployment_df, by =c("month", "year"))
```

This final dataset now contains the pols, snp, and unemployment datasets, grouped by month and year.

* The pols dataset contains information from January 1947 to June 2015 related to the number of national politicians who are democratic or republican at any given time. 
* The snp dataset contains the closing values of the S&P from January 1950 to July 2015. 
* The unemployment dataset contains the monthly unemployment percentages from January 1948 to June 2015. 
* The final dataset has a total of `r nrow(complete_df)` observations and `r ncol(complete_df)` variables. 
* The final dataset has the same date range as that of the pols dataset since we merged subsequent datasets into the pols dataset using left join. 
* The variables contained in the final dataset are: `r colnames(complete_df)`.
